---
title: "Heuristic"
format: html
editor: visual
---

```{r}
library(targets)
library(igraph)
library(tidyverse)
tar_config_set(
  script = here::here('_targets.R'),
  store = here::here('_targets')
)
theme_set(theme_classic())
source(here::here("R/functions.R"))
```

## Load data

```{r}
tar_load(timewindows)
tar_load(graphs_flight)
tar_load(graphs_feeding)
tar_load(graphs_roosting)
tar_load(graphs_aggregate)
```

General TWIN algorithm from the paper Meaningful Selection of Temporal Resolution for Dynamic Networks (Sulo et al. 2010)

for w = 1 to wmax, do:

compute the time series of graphs G(w)

Compute the time series Fw

if V(Fw) - R(Fw) \< gamma then

output w

end if

end for

Okay, so we have G(w) for each of five w values (stored above as graphs_flight).

Now let's compute F(w) for a few different measures. Density and diameter are easiest in igraph so let's start there, and we can also add transitivity

```{r}
nonzero <- function(g){
  gg <- delete_edges(g, which(E(g)$weight==0))
  return(gg)
}

fl_density <- map(graphs_flight, ~map_dbl(.x, ~igraph::edge_density(nonzero(.x))))
fl_trans <- map(graphs_flight, ~map_dbl(.x, ~igraph::transitivity(nonzero(.x))))
fl_diam <- map(graphs_flight, ~map_dbl(.x, ~igraph::diameter(nonzero(.x))))
fe_density <- map(graphs_feeding, ~map_dbl(.x, ~igraph::edge_density(nonzero(.x))))
fe_trans <- map(graphs_feeding, ~map_dbl(.x, ~igraph::transitivity(nonzero(.x))))
fe_diam <- map(graphs_feeding, ~map_dbl(.x, ~igraph::diameter(nonzero(.x))))
ro_density <- map(graphs_roosting, ~map_dbl(.x, ~igraph::edge_density(nonzero(.x))))
ro_trans <- map(graphs_roosting, ~map_dbl(.x, ~igraph::transitivity(nonzero(.x))))
ro_diam <- map(graphs_roosting, ~map_dbl(.x, ~igraph::diameter(nonzero(.x))))
ag_density <- map(graphs_aggregate, ~map_dbl(.x, ~igraph::edge_density(nonzero(.x))))
ag_trans <- map(graphs_aggregate, ~map_dbl(.x, ~igraph::transitivity(nonzero(.x))))
ag_diam <- map(graphs_aggregate, ~map_dbl(.x, ~igraph::diameter(nonzero(.x))))
```

Let's visualize those over time, just for the sake of it

```{r}
#| warning: false
#| message: false

getmeasures <- function(dens, trans, diam, tw, sitch){
  out <- pmap(list(dens, trans, diam, tw),
                    ~data.frame(dens = ..1, 
                                trans = ..2, 
                                diam = ..3, 
                                w = ..4, 
                                situ = sitch) %>%
                     mutate(step = 1:n(),
                            time = step*w))
  return(out)
}
measures_fl <- getmeasures(fl_density, fl_trans, fl_diam, 
                           timewindows, "flight")
measures_fe <- getmeasures(fe_density, fe_trans, fe_diam, 
                           timewindows, "feeding")
measures_ro <- getmeasures(ro_density, ro_trans, ro_diam, 
                           timewindows, "roosting")
measures_ag <- getmeasures(ag_density, ag_trans, ag_diam, 
                           timewindows, "aggregate")
all <- list(measures_fl, measures_fe, measures_ro, measures_ag)
all <- map(all, list_rbind)

measures <- do.call(rbind, all) %>%
  rename("density" = dens,
         "transitivity" = trans,
         "diameter" = diam) %>%
  pivot_longer(cols = c("density", "transitivity", "diameter"), names_to = "measure", values_to = "value")

# Visualize: daily
measures %>%
  ggplot(aes(x = w, y = value, col = situ))+
  facet_wrap(~situ*measure, ncol = 3, scales = "free_y")+
  geom_point(pch = 1, alpha = 0.2)+
  geom_smooth()+
  scale_color_manual(name = "Situation", values = c("firebrick3", situcolors))
  theme(legend.position = "none")
```

Now let's plot the CV (sd/mean)

```{r}
cvs <- measures %>%
  group_by(w, situ, measure) %>%
  summarize(mn = mean(value, na.rm = T),
            sd = sd(value, na.rm = T),
            cv = sd/mn) %>%
  ungroup() %>%
  mutate(w_days = w)

cvs %>%
  ggplot(aes(x = w_days, y = cv, col = situ))+
  facet_wrap(~measure, scales = "free_y")+
  geom_line()+
  geom_point()+
  ylab("CV")+
  xlab("Time window (days)")+
  scale_color_manual(name = "Situation", values = c("firebrick3", situcolors))+
  theme(legend.position = "bottom",
        text = element_text(size = 16))

## same thing with time on a log scale:
cvs %>%
  ggplot(aes(x = log(w_days), y = cv, col = situ))+
  facet_wrap(~measure, scales = "free_y")+
  geom_line()+
  geom_point()+
  ylab("CV")+
  xlab("Time window (days) (log-transformed)")+
  scale_color_manual(name = "Situation", values = c("firebrick3", situcolors))+
  theme(legend.position = "bottom",
        text = element_text(size = 16))

```
